---
title: "Workshop: Data visualization"
format: html
editor_options: 
  chunk_output_type: console
---


## Setup

These are the steps needed to start up the workshop.


```{r}

# Install exscidata
library(devtools)
install_github("dhammarstrom/exscidata")

# Additional packages that we will use

# To install these packages e.g., 
# install.package("ggtext")

library(tidyverse)
library(ggtext)
library(cowplot)



# Load the data 
dat <- exscidata::thorstensson

?thorstensson


```



## The basics

- Data
- Aesthetics
- Geometric objects
- Facets
- Groups
- Scales

```{r}

ggplot(data = dat, 
       aes(x = ft_area_pct, 
           y = peak_torque_pct_mvc, 
           color = group)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  
  scale_color_brewer(type = "div")
  
  
  



```



## Annotations

- Labels
- Annotations


```{r}



ggplot(data = dat, 
       aes(x = ft_area_pct, 
           y = peak_torque_pct_mvc, 
           color = group)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  
  labs(x = "Fast twitch fiber are", 
       y = "Relative peak torque", 
       color = "Group") +
  
  annotate("text", 
           label = "Downhill skiers", 
           x = 75, 
           y = 80) +
  annotate("point", 
           shape = 21, 
           size = 4,
           x = 90, y = 90) +
  annotate("segment", 
           x = 70, xend = 90, 
           y = 70, yend = 90, 
           arrow = arrow(type = "closed")) +
    annotate("curve",
             curvature = -2,
           x = 70, xend = 90, 
           y = 70, yend = 90, 
           arrow = arrow(type = "closed"))
  
  
  



```




## More scales, color, shapes etc.

- Control color, fill and shape
- Color manipulation via alpha


```{r}


ggplot(data = dat, 
       aes(x = ft_area_pct, 
           y = peak_torque_pct_mvc, 
           fill = group, 
           shape = group)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  
  labs(x = "Fast twitch fiber are", 
       y = "Relative peak torque", 
       fill = "Group", 
       shape = "Group") +
  
  scale_shape_manual(values = c(21, 22, 23, 24, 4))
  


```



## Themes

- Built in themes
- Changing theme components
- Building a custom theme

```{r}



my_theme <- function() {
    
  theme(legend.key.size = unit(6, "mm"), 
        axis.ticks.length = unit(10, "mm"))
  
}



ggplot(data = dat, 
       aes(x = ft_area_pct, 
           y = peak_torque_pct_mvc, 
           fill = group, 
           shape = group)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  
  labs(x = "Fast twitch fiber are", 
       y = "Relative peak torque", 
       fill = "Group", 
       shape = "Group") +
  
  scale_shape_manual(values = c(21, 22, 23, 24, 4)) +
  
  theme_classic() + 
  my_theme()
  
  
  
  


```


## More on annotations

- ggtext for markdown styled text


```{r}


ggplot(data = dat, 
       aes(x = ft_area_pct, 
           y = peak_torque_pct_mvc, 
           fill = group, 
           shape = group)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  
  labs(x = "Fast<sub>twitch</sub> fiber <sup>area</sup>, &rarr; &deg;C", 
       y = "Relative peak torque", 
       fill = "Group", 
       shape = "Group") +
  
  scale_shape_manual(values = c(21, 22, 23, 24, 4)) +
  
  theme_classic() +
  theme(axis.title.x = element_markdown())
  




```



## Combining multiple plots

- Using cowplot

```{r}

p1 <- ggplot(data = dat, 
       aes(x = ft_area_pct, 
           y = peak_torque_pct_mvc, 
           fill = group, 
           shape = group)) +
  geom_point() +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  
  labs(x = "Fast<sub>twitch</sub> fiber <sup>area</sup>, &rarr; &deg;C", 
       y = "Relative peak torque", 
       fill = "Group", 
       shape = "Group") +
  
  scale_shape_manual(values = c(21, 22, 23, 24, 4)) +
  
  theme_classic() +
  theme(axis.title.x = element_markdown())


p2 <- ggplot(data = dat, 
       aes(x = group, 
           y = ft_area_pct)) + 
  geom_boxplot()



plot_grid(p1, p2, 
          ncol = 1, 
          rel_heights = c(6, 4)) +
  annotate("text", 
           x = c(0.02,0.02), 
           y = c(0.98,0.38), 
           label = c("A", "B"))





```





## Saving figures 

- Saving to a suitable output format
- A workflow for developing figures for publication












